### 2015-05-27 doc/manual/inputParameters.tex
Added to user manual, better control of mumps cntl 1

@@

{\ttfamily geometryScheme}==12: Read the IPP {\ttfamily .bc} format Boozer-coordinate file specified in {\ttfamily equilibriumFile} below. 
The file is assumed to be stellarator-asymmetric.
}

@@

\param{inputRadialCoordinate}
{integer}
{3}
{When {\ttfamily geometryScheme} == 1, 5, 11, or 12}
{Which radial coordinate to use to specify the flux surface for a single calculation,
or to specify the range of flux surfaces for a radial scan. 
(When {\ttfamily geometryScheme} == 2, 3, or 4, the flux surface used will be {\ttfamily rN}=0.5.)
See section \ref{sec:radialCoordinates}
for more information about radial coordinates.\\

@@

This namelist defines which species are included in the calculation, along with the density and temperature and gradients thereof.
You will definitely want to set the parameters in this namelist.

>>>>

(All the variables in this formula are discussed later in this namelist.)\\

@@

{\ttfamily geometryScheme}==12: Read the IPP {\ttfamily .bc} format Boozer-coordinate file specified in {\ttfamily equilibriumFile} below. 
The file is assumed to be stellarator-\emph{asymmetric}.
}

@@

\param{inputRadialCoordinate}
{integer}
{3}
{When {\ttfamily geometryScheme} == 1, 5, 11, or 12}
{Which radial coordinate to use to specify the flux surface for a single calculation,
or to specify the range of flux surfaces for a radial scan. 
(Regardless of the value of this parameter, when {\ttfamily geometryScheme} == 2, 3, or 4, the flux surface used will be {\ttfamily rN}=0.5.)
See section \ref{sec:radialCoordinates}
for more information about radial coordinates.\\

@@

This namelist defines which species are included in the calculation, along with the density and temperature and gradients thereof.
You will definitely want to set the parameters in this namelist.
Note that only one of the four parameters 
{\ttfamily dnHatdpsiHats}, 
{\ttfamily dnHatdpsiNs}, 
{\ttfamily dnHatdrHats}, or
{\ttfamily dnHatdrNs}
will be used, depending on the value of {\ttfamily inputRadialCoordinateForGradients}
in the {\ttfamily geometryParameters} namelist.
Similarly, only one of the four parameters
{\ttfamily dTHatdpsiHats}, 
{\ttfamily dTHatdpsiNs}, 
{\ttfamily dTHatdrHats}, or
{\ttfamily dTHatdrNs}
will be used.

### 2015-05-27 doc/manual/installation.tex
Added to user manual, better control of mumps cntl 1

@@

Second, you will need to edit {\ttfamily utils/sfincsScan}. Look for the {\ttfamily if} block near the top
with sections for {\ttfamily sfincsSystem = edison},{\ttfamily hydra}, and {\ttfamily laptop}. Add an analogous block
for your system to set the command used to submit jobs, and a {\ttfamily nameJobFile} function.

@@

>>>>

To eliminate the need to set the environment variable and path 
as described above
at each login session, you may find it convenient to set both
in your startup script, such as  {\ttfamily .bashrc}.  In this startup script
you may also want to load any modules needed by \sfincsScan~such as \python, {\ttfamily numpy},
{\ttfamily scipy}, and {\ttfamily matplotlib} (if your computing system uses modules).

@@

Second, you will need to edit {\ttfamily utils/sfincsScan}. Look for the {\ttfamily if} block near the top
with sections for {\ttfamily sfincsSystem = edison}, {\ttfamily hydra}, and {\ttfamily laptop}. Add an analogous block
for your system to set the command used to submit jobs, and a {\ttfamily nameJobFile} function.

@@

The {\ttfamily make test} functionality relies on system-specific commands in the
\path{sfincs/fortran/version3/makefiles/makefile.SFINCS_SYSTEM} file.  If you experience
problems with  {\ttfamily make test}, it is likely that this file needs modification.

### 2015-05-27 doc/manual/overview.tex
Added to user manual, better control of mumps cntl 1

The \sfincs~code is a freely available, open-source tool for solving neoclassical-type kinetic problems in nonaxisymmetric or axisymmetric plasmas
with nested toroidal flux surfaces.
The code solves a drift-kinetic equation for an arbitrary number of species.  Optionally, a quasi-neutrality
equation can also be solved to obtain the self-consistent variation of the electrostatic potential on a flux surface.

This manual describes ``version 3'' of \sfincs.  There are two older versions of the
code called singleSpecies and multiSpecies.

@@

Often, the limiting factor for \sfincs~is the ability of \mumps~or \superludist~to factorize the preconditioner matrix.
You may therefore find it useful to see the control parameters and error codes in the \mumps~user manual:
\url{http://mumps.enseeiht.fr/doc/userguide_5.0.0.pdf}.

\section{Geometry options}
In \sfincs, a variety of options are available for the magnetic field geometry.  The geometry can be read directly
from a {\ttfamily vmec wout}
file, or from the {\ttfamily .bc} format Boozer-coordinate data files used at the Max Planck Institute for Plasma Physics (IPP).
A general analytic model for the magnetic field is also available, given by equation (\ref{eq:Bmodel}),
as are several analytic models for LHD and W7-X in which 3 or 4 Fourier components are retained.
For more details about geometry options in \sfincs, see section \ref{sec:geometryParameters}.

@@

\section{\sfincs~vs. \sfincsScan}
The core fortran part of \sfincs~solves the kinetic equation for each species
at a single flux surface, a single value of $E_r$, and a single set of other parameters.
However, often the goal is to determine the ambipolar $E_r$ at one or more surfaces, or to scan some other parameter.
For this task, the \sfincsScan~family of \python~scripts is available.
For a full list of the types of scans available, see section \ref{sec:sfincsScanParams}

@@

To report any bugs, provide feedback, or ask questions, contact Matt Landreman at
\href{mailto:matt.landreman@gmail.com}{\nolinkurl{matt.landreman@gmail.com} }

>>>>

The \sfincs~code is a freely available, open-source tool for solving neoclassical-type kinetic problems in nonaxisymmetric or axisymmetric plasmas
with nested toroidal flux surfaces.
As with other neoclassical codes, the input information used by \sfincs~is the equilibrium magnetic geometry together with the density, radial density gradient, temperature,
and radial temperature gradient of each species.  The code then solves a drift-kinetic equation for each species,
yielding the (gyro-angle averaged) distribution function.  Moments of the distribution function are computed such as the parallel
flow, bootstrap current, radial particle flux, radial heat flux, and variation of the density over a flux surface.  
These moments are all saved in the output file, and if you wish, you can also save the distribution function itself.
Optionally, a quasi-neutrality
equation can be solved at the same time as the drift-kinetic equations, yielding the self-consistent variation of the electrostatic potential on a flux surface.

The kinetic equations solved in \sfincs~have four independent variables: poloidal angle $\theta$, toroidal angle $\zeta$,
normalized speed $x = v / v_{thermal}$, and pitch angle $\xi = v_{||}/v$.  The third velocity coordinate (gyro-angle) does not appear
since gyro-averaged equations are solved.  The flux surface label (radius) coordinate is only a parameter, rather than a full independent variable,
since a radially local approximation is made.

@@

Often, the limiting factor for \sfincs~is the ability of the libraries \mumps~or \superludist~to factorize the preconditioner matrix, discussed in section \ref{sec:gmres}.
You may therefore find it useful to see the control parameters and error codes in the \mumps~user manual:
\url{http://mumps.enseeiht.fr/doc/userguide_5.0.0.pdf}.

This manual describes ``version 3'' of \sfincs.  
To preserve previous versions of the code that have been used for publications, two older versions of the
code called singleSpecies and multiSpecies are also present in the repository.  For both of these older versions,
both MATLAB and fortran editions exist which are independent of each other.  At present, version 3 exists
only in a fortran edition.

\section{Features}

\begin{itemize}

\item
Both self-species and inter-species collisions are treated using the most accurate linear operator available, the full linearized Fokker-Planck collision operator,
with no approximation of the field-particle term or expansion in mass ratio.  This collision operator conserves mass, momentum, and energy.

\item
Realistic experimental geometry can be simulated using an interface to \vmec.  Analytic model equilibria can also be used.

\item
Full coupling in the speed (or equivalently, kinetic energy) coordinate is retained, i.e. no monoenergetic approximation is made.  However,
if desired, \sfincs~can also be run in monoenergetic mode to compare with older codes.

\item
The code is formulated to permit solution of a wide variety of kinetic equations, whether or not phase space volume and/or energy are conserved, so individual terms can be turned on or off to examine their effect.

\item
A variety of models for terms involving the radial electric field are available to allow comparison between models.

\item
The code takes advantage of modern algorithms (GMRES) and parallelized libraries (\PETSc, \superludist, and \mumps).

\item
Efficient representation of velocity space is achieved using a pseudospectral method based upon non-classical orthogonal polynomials. \cite{speedGrids}

\item
Optional nonlinear terms in the kinetic equation (involving both the non-Maxwellian distribution function and poloidal/toroidal electric field) can be included using Newton's method.

\end{itemize}

\section{Limitations}

\begin{itemize}

\item
The \sfincs~code is radially local, in the sense that it approximates 
the radial derivative of the distribution function $\partial f/\partial \psi$ by
the derivative of a Maxwellian flux function $\partial f_M(\psi,x)/\partial \psi$.
This approximation is important for reducing the otherwise 5D space of independent variables $(\psi,\theta,\zeta,x,\xi)$
to a 4D space $(\theta,\zeta,x,\xi)$.
As a result, \sfincs~cannot compute certain finite-orbit-width effects that occur when the radial extent of the particle orbits
between bounces or transits is not small compared to the scale of radial variation in the equilibrium.
Such finite orbit width effects are significant near the magnetic axis, and in strong transport barriers, such as the pedestal of a tokamak H-mode.

\item
Turbulence is neglected.  There are good theoretical reasons to expect that the neoclassical effects computed by \sfincs~should
decouple from turbulence, as detailed in \cite{AbelReview}.  However, this argument relies on an expansion in $\rho_* \ll 1$, and so may break down
in some circumstances when $\rho_*$ is not sufficiently small.

\item
It is assumed that nested toroidal magnetic surfaces exist. Thus, the code cannot accurately model regions of stochastic field,
magnetic islands, or open field lines.

\end{itemize}

\section{Geometry options}
In \sfincs, a variety of options are available for the magnetic field geometry.  The geometry can be read directly
from a {\ttfamily vmec wout}
file, or from the {\ttfamily .bc} format Boozer-coordinate data files used at the Max Planck Institute for Plasma Physics (IPP).
A general analytic model for the magnetic field is also available, given by equation (\ref{eq:Bmodel}),
as are several analytic models for LHD and W7-X in which 3 or 4 Fourier components are retained.
The primary switch for controlling the magnetic geometry in \sfincs~is the {\ttfamily geometryScheme} parameter
in the {\ttfamily geometryParameters} input namelist.
For more details about geometry options in \sfincs, see section \ref{sec:geometryParameters}.

@@

\section{\sfincs~vs. \sfincsScan}
The core fortran part of \sfincs~solves the kinetic equation for each species
at a single flux surface, a single value of $E_r$, and a single set of other parameters.
However, often the goal is to determine the ambipolar $E_r$ at one or more surfaces, or to scan some other parameter.
For this task, the \sfincsScan~family of \python~scripts is available.
Using these scripts, it is also possible to scan other variables in the input file,
and in particular, to scan the resolution parameters to ensure
the physical output quantities are numerically converged.
For a full list of the types of scans available, see section \ref{sec:sfincsScanParams}

@@

We enthusiastically welcome any contributions to the code or documentation.
For write permission to the repository, or to report any bugs, provide feedback, or ask questions, contact Matt Landreman at
\href{mailto:matt.landreman@gmail.com}{\nolinkurl{matt.landreman@gmail.com} }
